---
title: 'Dynamic approach of spatial segregation: a framework with mobile phone data'
author: "Lino Galiana (INSEE)"
date: "November 15th, 2019"
institute: |
  | With Benjamin Sakarovitch (INSEE), François Sémécurbe (INSEE) and Zbigniew Smoreda (Orange Labs)
  |
  | JMA 2021 Virtual Conference
output:
  beamer_presentation:
    keep_tex: true
    toc: true
    slide_level: 3
    theme: "default"
    includes:
      in_header: header_pagenrs.tex
  keep_tex: true
  toc: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(tinytex.verbose = TRUE)
```


# Introduction

## Why do we need mobile phone data to measure segregation ? 

<!------------------------------------------------------->

### Residential segregation drivers: housing

* Income gradient from `r kableExtra::text_spec("housing prices", color = "darkblue")` (Alonso, `r kableExtra::text_spec("1964", color = "red")`)
    + High opportunity cost of transportation: wealthiest live in city center, poorest in suburbs
    + High valuation of housing space: wealthiest live in suburbs, poorest in city center
\medskip

* Social housing aims to ensure social mixing
    + Social housing clusters poor population in specific areas (Verdugo and Toma, `r kableExtra::text_spec("2018", color = "red")`)
    + Dynamic effect: school segregation creates persistence
    + People can coexist without interaction (Chamboredon and Lemaire, `r kableExtra::text_spec("1970", color = "red")`)


<!------------------------------------------------------->


<!------------------------------------------------------->

### Residential segregation drivers: preferences and mobility

* `r kableExtra::text_spec("Heterogeneity in preferences", color = "darkblue")` have spatial effects
    + Schelling (`r kableExtra::text_spec("1969", color = "red")`): clustering based on preference for neighborhood
    + Tiebout  (`r kableExtra::text_spec("1956", color = "red")`): spatial sorting based on public goods preferences
\medskip

* `r kableExtra::text_spec("Mobility", color = "darkblue")` plays a key role to understand segregation
    + Long run: high quality public good bring people in neighborhood, affecting housing price (Black, `r kableExtra::text_spec("1999", color = "red")`; Fack and Grenet, `r kableExtra::text_spec("2010", color = "red")`)
    + Within-week mobility brings together people from different neighborhood
\medskip

* `r kableExtra::text_spec("Infraday", color = "darkblue")` dynamic can be strong:
    + Davis et al. (`r kableExtra::text_spec("2019", color = "red")`): outside segregation (restaurants) 50\% lower than residential segregatio
    + Athey et al. (`r kableExtra::text_spec("2019", color = "red")`): similar scale for public space as parks




<!---
\item Some evidence on the effect of preferences over individual location
  \begin{itemize}
    \item Black (\textcolor{red}{1999}): parents are willing to pay 2.5 percent more for a 5 percent increase in test scores.
  \end{itemize}  \medskip
---->

<!------------------------------------------------------->

### Goal of the paper

From a picture

\begin{figure}
  \centering
  \begin{subfigure}[c]{0.45\textwidth}
      \centering
      \includegraphics[width=\linewidth]{./pics/income_D1_Paris_map.pdf}
      \caption{Low-income population (first decile)}
  \end{subfigure}\hskip 1em%
  \begin{subfigure}[c]{0.45\textwidth}
      \centering
      \includegraphics[width=\linewidth]{./pics/income_D9_Paris_map.pdf}
      \caption{High-income population (last decile)}
  \end{subfigure}
\end{figure}

to a more complete \href{https://github.com/linogaliana/prezJMAsegregation/blob/master/output/movie.gif}{sequence}

<!------------------------------------------------------->

<!------------------------------------------------------->

### Residential segregation: limitations of tax data

\begin{itemize}
  \item Good picture of residential segregation with tax \& census data
  \item But fixed picture
  \begin{itemize}
    \item People spend time out of their living neighborhood:
    \item Experienced segregation vs residential segregation
  \end{itemize}
  \item Numeric traces useful to know where people go
  \begin{itemize}
      \item Davis et al. (`r kableExtra::text_spec("2019", color = "red")`): Yelp data
      \item Athey et al. (`r kableExtra::text_spec("2020", color = "red")`): GPS data
  \end{itemize}
\end{itemize}

<!------------------------------------------------------->


<!------------------------------------------------------->

### Residential segregation: limitations of tax data

\begin{itemize}

  \item \textcolor{darkblue}{Theil index} (Theil, `r kableExtra::text_spec("1984", color = "red")`)
    \begin{align*}
      H = \frac{n_c}{N^{\text{city}}} \sum_{c=1}^C \frac{E(p^{\text{city}}) - E(p_c)}{E(p^{\text{city}})}
    \end{align*}
    \begin{itemize}
    \item Entropy measures diversity ($E(p) = -p\log(p) - (1-p)\log(1-p)$)
    \item Compares entropy at city and cell level
    \end{itemize}
  \item Administrative data $\Rightarrow$ \textcolor{darkblue}{residential segregation}:
  \begin{itemize}
    \item Static vision of segregation
    \item Separation of income groups within residential space
    \item No information on visited places
  \end{itemize} \medskip
  \item Mobility continuously reshapes income spatial distribution
  \begin{itemize}
    \item Need high-frequency geolocated data...
    \item ... combined with traditional data to characterize individuals
  \end{itemize}

\end{itemize}

<!------------------------------------------------------->


<!------------------------------------------------------->

## Research question

<!------------------------------------------------------->

### Research question

* Main questions:
    + How do mobility affect urban segregation ?
    + Do high-frequency data help us in identifying patterns in segregation that cannot be understood with administrative data?
    + Can we measure heterogeneity in spatial frictions within a city using high resolution mobility flows ?
\medskip

* Contribution:
    + Combining phone and traditional data
    + Proposition of a methodology to ensure combination robustness
    + Fine spatial and temporal granularity to understand segregation
    + Gravity approach with large scale data to measure cost of mobility



<!--------
\begin{itemize}
  \item Dissimilarity index (Duncan and Duncan, \textcolor{red}{1955} ; Massey and Denton, \textcolor{red}{1988}) is frequently used
  \begin{align*}
    ID = \frac{1}{2} \sum_{j=1}^J \bigg| \frac{w_j}{W_T}-\frac{n_j - w_j}{N_T - W_T} \bigg|
  \end{align*}
\end{itemize}
---->

<!------------------------------------------------------->


<!------------------------------------------------------->
### Methodology adopted

* We analyze `r kableExtra::text_spec("infraday dynamic", color = "darkblue")`:
    + `r kableExtra::text_spec("48 points", color = "darkblue")`: 24 for weekdays, 24 for weekend
\medskip

* Requires `r kableExtra::text_spec("time depending segregation indexes", color = "darkblue")`
    + Theil index series for each city
\medskip

* Gravity model to measure spatial frictions
    + Takes into account the zero-flows problem
\medskip

* `r kableExtra::text_spec("Paris, Lyon and Marseille", color = "darkblue")`
    + Agglomeration level: city centers and suburbs
    + More than 13 millions people in tax data

<!------------------------------------------------------->

# Data

### Principle

* Caracterize phone users from living environment
* Probability of belonging to first/last decile from observed income distribution in tax data

\includegraphics[page=12, width = \linewidth]{./pics/CNIS.pdf}



## Phone data

<!------------------------------------------------------->
### Phone data {label="phone data"}


* `r kableExtra::text_spec("Orange data September 2007", color = "darkblue")`
    + `r kableExtra::text_spec("18.5 millions", color = "darkblue")` SIM cards ($\approx$ 1/3 French population)
    + Text messages and call: 3 billions events
    + `r kableExtra::text_spec("Geocoding at antenna level", color = "darkblue")` (exact $(x,y)$ unknown)
\medskip

* Transformation into 500x500 meters cell level presence
\hyperlink{probabilization}{\beamerbutton{Methodology here}}
\medskip

* We do not use interaction dimension
    + Plan for future research on social segregation
\medskip

* Big data volume is a challenge


<!------------------------------------------------------->


<!------------------------------------------------------->
### Phone data

\begin{itemize}
  \item 2007 is old:
\begin{itemize}
  \item People were not using their phone as much as now
  \item Temporal sparsity at individual level (in average 4 points a day by user)
\end{itemize}
\end{itemize}


\begin{tiny}
\begin{table}[]
    \centering
    \begin{tabular}{p{0.3\linewidth}ccccccccc}
\toprule
& mean & s.d. & min & P10 & P25 & median & P75 & P90 & max  \\
\midrule
Average number of daily events per user & $4.3$ & $3.6$ & $1$ & $1.4$ & $2$ & $3.1$ & $5.4$ & $8.7$ & $123$ \\
Number of distincts days users appear & $20$ & $9.2$ & $1$ & $5$ & $13$ & $23$ & $28$ & $30$ & $30$ \\
Average number of events between 7PM and 9AM per user & $2.4$ & $1.7$ & $0$ & $1$ & $1.3$ & $1.9$ & $2.9$ & $4.4$ & $87$ \\
Number of distincts days users appear between 7PM and 9AM & $15.2$ & $9.4$ & $0$ & $2$ & $7$ & $15$ & $24$ & $28$ & $30$ \\
\midrule
\multicolumn{4}{l}{Number of observations: } & \multicolumn{2}{c}{3,024,884,663} \\
\multicolumn{4}{l}{Number of unique phone users:} & \multicolumn{2}{c}{18,541,440} \\
\bottomrule
    \end{tabular}
    \caption{Orange 2007 CDR : summary statistics of September data}
    \label{tab:statCDR2007}
\end{table}

\end{tiny}

<!------------------------------------------------------->


## Tax data




### Tax data

* `r kableExtra::text_spec("2014 geocoded tax data", color = "darkblue")` at $(x,y)$ level
    + Income by consumption unit
\medskip

* Income based segregation
    + Distribution of income extremes (first and last deciles)
    + Relative definition of income: is individual wealthier/poorer than a city reference level ?
\medskip

*  `r kableExtra::text_spec("Bimodal approach", color = "darkblue")`
    + First decile vs others
    + Last  decile vs others
\medskip

### Tax data

\begin{itemize}
  \setlength\itemsep{0.5em}
  \item Sub-population (first/last decile) frequency in cell
  \item \textcolor{darkblue}{Spatial aggregation} at cell level $i$
  \begin{align*}
      p_i^{D1} = \mathbb{P}\big(y_x < \mu^{D1}  \big) = \mathbb{E}\big( \mathbf{1}_{\{y_x < \mu^{D1}\}} \big) = \frac{1}{n_i} \sum_{x = 1}^{n_i} \mathbf{1}_{\{y_x < \mu^{D1}\}} \\
      p_i^{D9} = \mathbb{P}\big(y_x > \mu^{D9}  \big) = \mathbb{E}\big( \mathbf{1}_{\{y_x > \mu^{D9}\}} \big) = \frac{1}{n_i} \sum_{x = 1}^{n_i} \mathbf{1}_{\{y_x > \mu^{D9}\}}
  \end{align*}
  \item If $p_i>0.1$, over-representation of subpopulation in cell
  \item That frequency is used to simulate phone user status given their simulated residence
\end{itemize}


### Tax data

* Intuitions regarding city segregation from tax data
    + e.g. Paris: more segregation at the top

\begin{figure}[ht]
\centering
\includegraphics[width=0.6\textwidth]{./pics/dissimilarity_D1D9.pdf}
\caption{Dissimilarity index for main French cities}
\end{figure}


# Dynamic segregation

## Methodology to build segregation index

### Workflow {label="workflow methodo"}

* Phone user status is simulated from his/her phone track (only personal information) and neighborhood level tax aggregates
* 3 steps to estimate segregation dynamics:
    1. `r kableExtra::text_spec("Home estimation", color = "darkblue")`
        + Estimate probabilities that individual lives in some neighborhood given nighttime (19 pm - 9 am) phone track
    2. `r kableExtra::text_spec("Home cell and income simulations", color = "darkblue")`
        + Home simulation knowing cell level probability sequences
        + Income simulation given first/last decile frequence appearance in tax data ($p_i$)
        + Test other designs to check robustness of income simulation
    3. `r kableExtra::text_spec("Compute segregation indexes", color = "darkblue")`
        + They depend on observation time $t$ (dynamic approach)

\hyperlink{methodo home}{\beamerbutton{Details for step 1 and 2 here}}


### Segregation index {label="index prez"}

\begin{itemize}
\item Two typical days: weekdays, weekend
\item Individual probabilities at cell level on a given time window: $\mathbb{P}_x(c_{it})$ \hyperlink{annexeCarreaux}{\beamerbutton{Details}}
\item Probabilize \textcolor{darkblue}{Theil index} (as well as other indices):

\begin{small}
\begin{align*}
\label{eq: generalized theil index H}
    H_t = \sum_{c=1}^C
    \frac{
    \overbrace{\sum_{x \in \mathcal{X}} \mathbb{P}_{x,t}(c)}^{\substack{\text{Number people of income group }g \\ \text{that are observed at time } t \text{in cell } c}}
    }{
    \underbrace{\text{card}(\mathcal{X})}_{\substack{\text{Number people of income group }g \\ \text{that are observed at time } t}}
    }
    \frac{E(p_{c,t}) - E(p_{t}^{\text{city}})}{E(p_{t}^{\text{city}})}
\end{align*}
\end{small}

\item Remainder, standard index:
    \begin{align*}
      H = \frac{n_c}{N^{\text{city}}} \sum_{c=1}^C \frac{E(p^{\text{city}}) - E(p_c)}{E(p^{\text{city}})}
    \end{align*}
\end{itemize}


## Results at city level

### Segregation dynamics {label="segregation results"}

\begin{itemize}
\item City-level segregation evolution across time
  \begin{itemize}
  \item People not observed at a given hour of the night (19-9) are assumed to be at home
  \item This removes downward bias in index with respect to tax data
  \end{itemize}
  \item Dynamic robust to other income simulation methods
\begin{itemize}
  \item Alternative simulation: nighttime level affected but dynamics keep the same pattern
\end{itemize}
\end{itemize}

\begin{table}[]
\resizebox{\textwidth}{!}{%

  \centering
  \begin{tabular}{p{0.3\linewidth}cccccc}
  \toprule
  & \multicolumn{2}{c}{Paris} & \multicolumn{2}{c}{Lyon} & \multicolumn{2}{c}{Marseille} \\
  & Low-income & High-income & Low-income & High-income & Low-income & High-income \\
  \midrule
  & \multicolumn{6}{c}{Weekdays} \\
  Max amplitude & 0.05 & 0.12 & 0.06 & 0.1 & 0.08 & 0.11\\
  Relative amplitude (\%) & \textcolor{blue}{76.72} & 68.77 & 88.03 & 82.03 & 77.82 & 77.8 \\
  Within night (19h-9h) relative amplitude (\%) & \textcolor{blue}{61.67} & 55.5 & 71.46 & 65.56 & 64.46 & 61.48 \\
  \bottomrule
  \multicolumn{6}{l}{Max amplitude $H^{\text{max}} - H^{\text{min}}$} \\
  \multicolumn{6}{l}{Relative amplitude (\%): $1 - H^{\text{min}}/H^{\text{max}}$} \\
  \end{tabular}}
\end{table}


### Segregation dynamics: low-income

\begin{figure}[ht]
\centering
\includegraphics[width=0.9\textwidth]{./pics/bootstrap_index_H_D1.pdf}
\caption{Low-income segregation dynamics (50 replications bootstrap)}
\end{figure}



### Segregation dynamics: high-income

\begin{figure}[ht]
\centering
\includegraphics[width=0.9\textwidth]{./pics/bootstrap_index_H_D9.pdf}
\caption{High-income segregation dynamics (50 replications bootstrap)}
\end{figure}


### Segregation dynamics: comparing cities and income groups

* Significant difference between nighttime and daytime segregation levels
    + Segregation starts to decrease around 6-7am and goes up after 4-5pm
    + No significant difference between weekend and weekdays $\Rightarrow$ separate saturday and sunday ?
\medskip

* Differences in level observed in tax data also present in phone data
    + e.g. Paris: segregation higher at the top
\medskip

* Mobile phone inform us on dynamics:
    + Decrease stronger in Marseille and Lyon than in Paris
    + Track neighborhood composition \hyperlink{clustering}{\beamerbutton{Results here}}
\medskip
    + Further research: can we identify some inclusive/exclusive cities ?

### Evolution of city structure across time

\begin{small}
e.g. Low-income concentration at two different times (6am and 4pm). 
\href{https://github.com/linogaliana/prezJMAsegregation/blob/master/output/movie.gif}{Full sequence here}
\end{small}



\begin{figure}

  \centering
  \begin{subfigure}[c]{0.48\textwidth}
      \centering
      \includegraphics[width=\linewidth]{./pics/paris_mobile_hour6_D1_english.png}
  \end{subfigure}
  \begin{subfigure}[c]{0.48\textwidth}
      \centering
      \includegraphics[width=\linewidth]{./pics/paris_mobile_hour16_D1_english.png}
  \end{subfigure}


\end{figure}


## Clustering Paris by neighborhood population pattern

### Principle {label=clustering}

\begin{itemize}
  \item Clustering to identify spaces that share common population composition characteristics
  \begin{itemize}
    \item Will be related to places characteristics (infrastructures...)
  \end{itemize}
  \item sequence of low-income and high-income concentrations in our cells
  \item k-means classification algorithm
  \item 4 clusters is motivated by the trade-off between variety and parsimony.
\end{itemize}


### Results

\begin{figure}

  \centering
  \begin{subfigure}[c]{0.48\textwidth}
      \centering
      \includegraphics[width=\linewidth]{./pics/typo.png}
  \end{subfigure}
  \begin{subfigure}[c]{0.48\textwidth}
      \centering
      \includegraphics[width=\linewidth]{./pics/paris_kmeans_prof.pdf}
  \end{subfigure}


\end{figure}



# Gravity model from urban flows

## Specification

### Gravity model with origin-destination flows

\begin{align}
p_{i \to j}^g = a \frac{M_i^{\beta_1}M_j^{\beta_2}}{D_{ij}^{\beta_3}}
\end{align}


* Mobile phone literature refer to gravity equation (e.g. Krings *et al*, `r kableExtra::text_spec("2009", color = "red")`)
    + Does not estimate distance-decay with robust methodology
    + Some common caveats of gravity equation (e.g. zero-flows problem) need to be accounted
\medskip

* We observe only strictly positive flows (censoring problem)
    + Loglinearized OLS equations are biased
\medskip

* Silva \& Tenreyro (`r kableExtra::text_spec("2006", color = "red")`) and Silva \& Tenreyro (`r kableExtra::text_spec("2011", color = "red")`):
    + Augment observed sample with every potential flows
    + Count data models (Poisson) more suited than a log-linearized OLS equation
\medskip

* When large share of zeros (our case): zero-inflated count model

### Gravity model with origin-destination flows

* We propose to use estimation strategies derived from international trade theory...
\medskip

* ... with urban flows measured using mobile phone data
    + Likelihood of being in cell $c_i$ knowing people live in cell $c_j$
    + Origin-destination flows at 500 meters level
\medskip

* Estimate heterogeneity in distance costs:
    + Spatial dimension: suburbs vs center
    + Social dimension: low-income vs high-income


\begin{align}
\begin{array}{c @{} c}
& p_{i \to j}^g \sim \pi \delta_0 + (1-\pi) \mathcal{NB}(\lambda) \\
\text{(Selection) \qquad } &  \text{logit}(\pi) = \sum_i \alpha_i z_i + u_i \\
\text{(Outcome)\qquad} & \lambda(X_{ij}) = \mathbb{E}_{f, \theta}(p_{i \to j}^g|X_{ij}) = \exp(\beta^{\intercal} X_{ij}) \\ \end{array}
\end{align}

* BIC: Choice of negative binomial (outcome) with probit link (selection)

## Results

### Results (Marseille)


\input{tables/gravity/gravity_negbinom_marseille_subset.tex}

### Results (Paris)

\input{tables/gravity/gravity_negbinom_paris_subset.tex}


### Results (Lyon)

\input{tables/gravity/gravity_negbinom_lyon_subset.tex}


# Conclusion

### Conclusion

* Bringing together phone and tax data requires methodological foundations
* Segregation:
    + Acme during nighttime/hometime
    + Goes down by $\approx$ 50\% by daytime
    + Results consistent with Davis et al (`r kableExtra::text_spec("2019", color = "red")`) and Athey et al (`r kableExtra::text_spec("2019", color = "red")`)
* Mobility cost:
    + Depends on urban structure: Marseille vs Paris/Lyon
    + Some heterogeneity given neighborhood income level: e.g. low-income neighborhood in Marseille


<!------------------------- APPENDIX ----------------------------->
# Appendix

<!-------------
\begin{frame}[label=probabilization]{FDFF}

\hyperlink{phone data}{\beamerbutton{Methodology here}}.


\end{frame}
----------------->

## Probabilization

### Phone users' presence probabilization {label="probabilization"}

\hyperlink{phone data}{\beamerbutton{Back to slide}}

\begin{figure}[ht]
\begin{minipage}[b]{0.3\linewidth}
\centering
  \includegraphics[width = \textwidth]{./pics/points.pdf}
\end{minipage}
\hspace{0.5cm}
\begin{minipage}[b]{0.3\linewidth}
\centering
\includegraphics[width=\textwidth]{./pics/points_and_voronoi.pdf}
\end{minipage}
\end{figure}


\begin{itemize}
    \setlength\itemsep{1em}
  \item Mobile phone litterature does not dissociate:
  \begin{itemize}
  \item \textcolor{darkblue}{Coverage area}: observations at antenna level into presence area
  \item \textcolor{darkblue}{Statistical unit}: economic information level
  \end{itemize}
  \item Coverage area: \textcolor{darkblue}{Voronoi tesselation}
  \begin{itemize}
  \item Each point in space is associated with closest antenna
  \end{itemize}
  \item However, \textcolor{darkblue}{must not be analysis statistical unit}
  \begin{itemize}
  \item Partition depends too much on antennas local density
  \end{itemize}
\end{itemize}

<!---------------------------------------->

### Phone users' presence probabilization

```{r}
# source("illustration_voro_to_grid.R")
```


\begin{itemize}
  \item Cell level probabilization to abstract from voronoi
  \begin{itemize}
    \item Knowing call has been observed from antenna $v_j$, probability it happened into cell $c_i$? (Bayes rule)
  \end{itemize}
%\begin{align*}
%p_i^j := \mathbb{P}\big( c_i | v_j \big) = \frac{\mathbb{P}\big( c_i \cap v_j \big)}{\mathbb{P}(v_j)} = \frac{\mathcal{S}\big( c_i \cap v_j \big)}{\mathcal{S}(v_j)}
%\end{align*}
%\item Probabilités au niveau des carreaux $c_i$ sont toujours conditionnelles à des observations au niveau des antennes/voronoi $v_j$
\item \textcolor{darkblue}{500x500m cell level}
  \begin{itemize}
    \item Phone data: probabilize both presence and home
    \item Tax data: local aggregates at cell level
  \end{itemize}
  \item Illustration in next slide for home detection:
  \begin{itemize}
    \item 2/3 events located in $v_2$ ; 1/3 located in $v_1$
    \item Grid probabilities $(\mathbb{P}(c_i|v_j))_{i,j}$ via Bayes' rule (see (c) and (d))
    \item With uninformative prior, home detection given by (e)
    \item If population denser in tiles that intersect $v_1$ (f), home detection is modified (g)
  \end{itemize}



\end{itemize}

<!---------------------------------------->

<!---------------------------------------->

### Phone users' presence probabilization

\begin{figure}[ht]
    \centering

\begin{minipage}[t]{0.2\linewidth}
    \centering
    \includegraphics[width = \textwidth]{./pics/voro_grid_example/voro1.png}
\end{minipage} %
\begin{minipage}[t]{0.2\textwidth}
    \centering
    \includegraphics[width = \textwidth]{./pics/voro_grid_example/voro2.png}
\end{minipage} %

\begin{minipage}[t]{0.25\textwidth}
    \centering
    \includegraphics[width = \textwidth]{./pics/voro_grid_example/grid1.png}
\end{minipage} %
\begin{minipage}[t]{0.25\textwidth}
    \centering
    \includegraphics[width = \textwidth]{./pics/voro_grid_example/grid2.png}
\end{minipage} %
\begin{minipage}[t]{0.25\textwidth}
  \centering
  \includegraphics[width = \textwidth]{./pics/voro_grid_example/grid3.png}
\end{minipage}

\begin{minipage}[t]{0.25\textwidth}
  \centering
  \includegraphics[width = \textwidth]{./pics/voro_grid_example/prior1.png}
\end{minipage} %
\begin{minipage}[t]{0.25\textwidth}
  \centering
  \includegraphics[width = \textwidth]{./pics/voro_grid_example/prior2.png}
\end{minipage}

\end{figure}


<!---------------------------------------->



## Methodology: more details

<!---------------------------------------->
### 1. Home estimation {label="methodo home"}

\begin{itemize}
  \item Nighttime phone track (19h-9h) used to estimate individual residence probability for all cells
  \item Bayesian approach to account for the fact that all metropolitan space is not residential
  \begin{itemize}
    \item In a coverage area, prior in most densily populated cells
    \item Prior from population density computed from tax data
  \end{itemize}
  \item Prior distribution is a reweighting for cell level home
  \begin{align*}
    \mathbb{P}_x\big(c_i^{\text{home}}|v_j\big) \propto \underbrace{\mathbb{P}\big(c_i^\text{home}\big)}_{\substack{\text{prior from} \\ \text{population density}}}\underbrace{\mathbb{P}_x\big(v_j|c_i\big)}_{\substack{\text{ areas ratio:} \\ \frac{s(v \cap c)}{s(c)}}}
  \end{align*}
  \item Sequence from home probabilities: $\nu_x^{\text{home}}(c_i)$
  \begin{itemize}
    \item Used to simulate $x$ income
  \end{itemize}
\end{itemize}

\hyperlink{workflow methodo}{\beamerbutton{Back to presentation}}


### 2. Home and income simulations

\hfill

\textcolor{darkblue}{4 methods} of home simulation to check robustness of segregation indexes


\begin{small}
\begin{table}[]
\begin{tabular}{p{0.3\linewidth}p{0.7\textwidth}l}
\toprule
\multicolumn{1}{c}{Methodology} & \multicolumn{1}{c}{Choice of $x$'s home} \\
\midrule
\textcolor{darkblue}{Main method} & Draw home from all residence probabilities $\nu_x^{\texttt{home}}$  \\
\textcolor{darkblue}{\texttt{One stage simulation}} &  Cell where probability is maximum: $c_i = \arg \max_{c_i} \nu_x^{\text{home}}(c_i)$ \\
\textcolor{darkblue}{\texttt{cell\_max\_proba}} & $x$ assigned where probability of being member of group $g$ is maximized\\
\textcolor{darkblue}{\texttt{cell\_min\_proba}} & $x$ assigned where probability of being member of group $g$ is minimized \\
\bottomrule
\end{tabular}
\end{table}
\end{small}

Last two methods: evaluate effect on segregation indexes to over- or under-estimate the share of sub-group $g$ on population
\hyperlink{workflow methodo}{\beamerbutton{Back to presentation}}


<!---------------------------------------->


### 3. Segregation indexes: cell level presence {label=annexeCarreaux}


\begin{itemize}
\item Probability that an event measured in antenna $v_j$ at time $t$ occurred in cell $c_i$ is
\begin{align*}
p_i^j := \mathbb{P}\big( c_i | v_j \big) = \frac{\mathbb{P}\big( c_i \cap v_j \big)}{\mathbb{P}(v_j)} = \frac{\mathcal{S}\big( c_i \cap v_j \big)}{\mathcal{S}(v_j)}
\end{align*}
\item We denote $c_{it}$ the probability of being present at time $t$ in cell $c_i$. This is a recollection of conditional probabilities
\begin{align}
\label{eq: p(c)}
\forall c_{it} \in \mathcal{C}, \quad \mathbb{P}_x(c_{it}) = \sum_{v_{jt} \in \mathcal{V}} \mathbb{P}\big(c_{it} \big| v_{jt}\big)\mathbb{P}_x(v_{jt})
\end{align}
with $\mathcal{V}$ voronoi/antennas and $\mathcal{C}$ 500m cells.
\end{itemize}

\hyperlink{index prez}{\beamerbutton{Back to presentation}}

## Additional elements:  spatial clustering

### Additional elements:  spatial clustering  {label=clustering}

\hyperlink{segregation results}{\beamerbutton{Back to slide}}


\begin{itemize}
  \item Clustering to identify spaces that share common population composition characteristics
  \begin{itemize}
    \item Will be related to places characteristics (infrastructures...)
  \end{itemize}
  \item e.g.: share of population belonging to low-income group
\end{itemize}


